---
title: "Demographic and Social Stressor Data Preprocessing"
author: "Elise Hickman"
output:
  html_document:
    code_folding: show
    toc: true
    toc_depth: 6
    number_sections: false
---

# Script Summary

Script within this file was used to:

+ Clean demographics data
+ Generate the demographics table (Table S2)
+ Clean and imput social stressor data
+ Identify outliers in social stressor data (Figure S3)

# Set up workspace

```{r message = FALSE, warning = FALSE}
# Clear global environment
rm(list=ls())

# Load packages
library(tidyverse) # for data organization and manipulation
library(janitor) # for data cleaning
library(openxlsx) # for writing out files
library(DT) # for displaying tables
library(table1) # for making tables
library(factoextra) # for PCA outlier detection
library(missForest) # for imputation
library(patchwork) # for graphing
select <- dplyr::select
rename <- dplyr::rename

# Set theme
theme_set(theme_bw())

# Set working directory
setwd("/Users/ehickman/Library/CloudStorage/OneDrive-UniversityofNorthCarolinaatChapelHill/Rager_Lab/Projects/2023/BEE_Wristband_Project/4_Analyses/1_CurrentAnalyses/1_DataPreprocess/2_DemographicAndSSData")
```

# Import data

For this and following analyses, ss_data will refer to social stressor data. Initially, this also includes demographic information, but that will be filtered out below.

```{r}
ss_data <- data.frame(read.csv(file = "1_InputData/NonChemicalData.csv")) %>%
  clean_names() %>%
  rename("S_ID" = "s_id")
```

We will import the chemical data just to extract the participants that we want to keep/focus on. 
```{r}
# Chemical data
wb_data <- read.xlsx("1_InputData/ChemicalData_Filtered_Imp_TW.xlsx")

# Extract participant IDs
participant_IDs <- wb_data$S_ID
```

# Clean demographic data

Demographic columns that contained duplicate information were removed and sex, race, and ethnicity data were recoded to be more descriptive and streamlined. For example, the degree category was collapsed based on the highest degree completed (e.g., Bachelor's Degree Plus Additional Training was changed to Bachelor's Degree), and race/ethnicity categories were collapsed such that any indication of biracial overrode other indicated race, and Latino and Hispanic were combined into one category. 

```{r}
# Data cleaning
demo <- ss_data %>%
  filter(S_ID %in% participant_IDs) %>%
  select(S_ID:pc_ed) %>%
  select(-c(tc_age_days_t18, ga_days_t18, pc_matage_t18, 
            remote_t6, tc_age_days_t6, ga_days_t6, pc_matage_t6, tc_age_months_t18)) %>%
  rename("child_age_months" = "tc_age_months_t6",
         "mat_age_birth" = "pc_mat_agebirth",
         "child_gest_age_birth" = "ga_days_birth") %>%
  mutate(across(tc_female:pc_ed, as.character)) %>%
  mutate(tc_female = recode(tc_female, "0" = "M", "1" = "F")) %>%
  rename("child_sex" = "tc_female") %>%
  mutate(pc_sex = recode(pc_sex, "0" = "F", "1" = "M"),
         pc_gender = recode(pc_gender, "0" = "Woman", "1" = "Man", "2" = "Trans", "3" = "Genderqueer", "4" = "Other"),
         pc_latino = recode(pc_latino, "0" = "No", "1" = "MexicanChicano", "2" = "PuertoRican", "3" = "Cuban", "4" = "Other"),
         pc_latino = case_when(pc_latino == "No" ~ "No", .default = "Yes"),
         pc_race = recode(pc_race, "0" = "White", "1" = "Black", "2" = "IndigenousAmerican", 
                          "3" = "AsianIndian", "4" = "Chinese", "5" = "Japanese", "6" = "Korean", 
                          "7" = "Vietnamese", "8" = "OtherAsian", "9" = "Other"),
         pc_biracial = recode(pc_biracial, "0" = "No", "1" = "Yes"),
         pc_addrace = recode(pc_addrace, "0" = "White", "1" = "Black", "2" = "IndigenousAmerican", 
                             "3" = "AsianIndian", "4" = "Chinese", "5" = "Japanese", "6" = "Korean", 
                             "7" = "Vietnamese", "8" = "OtherAsian", "9" = "Other"),
         pc_race_black = recode(pc_race_black, "0" = "No", "1" = "Yes"),
         pc_ed = recode(pc_ed, "12" = "High School or GED", "13" = "High School or GED", "14" = "Associate Degree", 
                        "15" = "Associate Degree", "16" = "Four-Year Degree", "17" = "Four-Year Degree", 
                        "18" = "Master's Degree", "19" = "Professional Degree or PhD")) %>%
  select(-c(pc_race_black)) %>%
  mutate(pc_latino_hispanic = case_when(pc_latino == "Yes" ~ "Yes",
                                        pc_race_other == "Hispanic" ~ "Yes",
                                        pc_race_other == "Medican" ~ "Yes",
                                        pc_race_other == "Latina" ~ "Yes",
                                        .default = "No"),
         pc_multiracial_cleaned = case_when(pc_race_other == "biracial" ~ "Yes",
                                            pc_race_other == "multi" ~ "Yes",
                                            pc_biracial == "Yes" ~ "Yes",
                                            .default = "No"),
         pc_race_cleaned = case_when(pc_multiracial_cleaned == "Yes" ~ "Biracial/Multiracial",
                                     pc_race == "White" ~ "White",
                                     pc_race == "Black" ~ "Black",
                                     pc_race == "Chinese" ~ "Other",
                                     pc_race == "Other" ~ "Other")) %>%
  select(-c(pc_latino:pc_race_other, pc_multiracial_cleaned))
```

Relationship status data needed different cleaning because if the participant was in the same romantic relationship as they were in at the prenatal visit, many of the questions are skipped.
```{r}
# Filter for just relationship data
relationship_demo <- ss_data %>%
  filter(S_ID %in% participant_IDs) %>%
  select(c(S_ID, starts_with("rel")))

# Filter to subjects who are not in a relationship
relationship_demo_no <- relationship_demo %>%
  filter(rel1_t6 == 0) %>%
  mutate(relationship = "None") %>%
  select(c(S_ID, relationship)) %>%
  mutate("Total Years with Partner" = NA)

relationship_demo_no_IDs <- relationship_demo_no$S_ID 

# Filter to subjects who indicated they were in a relationship
relationship_demo_yes <- relationship_demo %>%
  filter(!S_ID %in% relationship_demo_no_IDs)

# Filter to subjects who indicated they were in a relationship and this has not changed since t0  (changing NAs to 3 so they will be treated properly)
relationship_demo_yes_same <- relationship_demo_yes %>%
   mutate(rel_new_t6 = replace_na(rel_new_t6, 3)) %>%
  filter(rel_new_t6 == 0 | rel_new_t6 == 3) %>%
  mutate(relationship = ifelse(rel2_t0 == 1, "Yes and Live Together", "Yes and Do Not Live Together")) %>%
  mutate(relationship = replace_na(relationship, "Yes (Unclear Living Status)")) %>%
  relocate(relationship, .after = "S_ID") %>%
  select(c(S_ID, relationship, rel_ttl_yrs_wpartner_t0)) %>%
  rename("Total Years with Partner" = "rel_ttl_yrs_wpartner_t0")

# Filter to subjects who indicated that their relationship has changed since t0
relationship_demo_yes_diff <- relationship_demo_yes %>%
  filter(rel_new_t6 == 1) %>%
  mutate(relationship = ifelse(rel2_t0 == 1, "Yes and Live Together", "Yes and Do Not Live Together")) %>%
  mutate(relationship = replace_na(relationship, "Yes (Unclear Living Status)")) %>%
  relocate(relationship, .after = "S_ID") %>%
  select(c(S_ID, relationship, rel_ttl_yrs_wpartner_t6)) %>%
  rename("Total Years with Partner" = "rel_ttl_yrs_wpartner_t6")

# Combine back data frames
relationship_demo_cleaned <- bind_rows(relationship_demo_no, relationship_demo_yes_diff, relationship_demo_yes_same)

# Combine relationship data with existing data
demo <- left_join(demo, relationship_demo_cleaned, by = "S_ID") %>%
  rename("yrs_with_part" = "Total Years with Partner")

# Re-level variables
demo$pc_race_cleaned <- factor(demo$pc_race_cleaned, 
                               levels=c("White", "Black", "Biracial/Multiracial", "Other"))

demo$pc_latino_hispanic <- factor(demo$pc_latino_hispanic, 
                               levels=c("No", "Yes"))

demo$pc_ed <- factor(demo$pc_ed, 
                               levels=c("High School or GED", "Associate Degree", "Four-Year Degree", 
                                        "Master's Degree","Professional Degree or PhD"))


demo$relationship <- factor(demo$relationship, 
                               levels=c("Yes and Live Together", "Yes and Do Not Live Together", "Yes (Unclear Living Status)", 
                                        "None"))

datatable(demo)
```

Write out data:
```{r}
write.xlsx(demo, "2_OutputData/Demographic_Data_Cleaned.xlsx")
```


# Make demographics table

Return to making demographics table. Will manually change "Missing" under years together to only reflect those who are in a relationship (subtract the 7 who are not in a relationship).

```{r}
# Create new labels for the demographics table
label(demo$mat_age_birth) <- "Age at Childbirth"
label(demo$pc_latino_hispanic) <- "Latino or Hispanic"
label(demo$pc_race_cleaned) <- "Race"
label(demo$pc_ed) <- "Educational Attainment"
label(demo$relationship) <- "Relationship Status"
label(demo$yrs_with_part) <- "Years with Partner"

# Create function for custom table so that Mean (SD) is shown for continuous variables
my.render.cont <- function(x) {
  with(stats.apply.rounding(stats.default(x), digits=2), c("",
                                                           "Mean (SD)"=sprintf("%s (&plusmn; %s)", MEAN, SD)))
}

# Make table of demographics.
table1(~ mat_age_birth + pc_race_cleaned + pc_latino_hispanic + pc_ed + relationship + yrs_with_part, 
       data = demo, 
       render.continuous = my.render.cont)
```

# Cleaning Social Stressor Data

## Organize data for imputation

```{r}
# Remove demographic data and select only subjects in the 6-month cohort
ss_data <- ss_data %>%
  select(-c(tc_age_months_t18:rel1_t0)) %>%
  filter(S_ID %in% participant_IDs) %>%
  mutate(S_ID = as.character(S_ID))

# Split out data by time point and clean data.

# HC data will be treated differently, as data are not missing at random. Participants with shorter and more kinky hair are more likley to have missing data here, so this will be approached as a subset analysis.
ss_data_hc <- ss_data %>%
  select(c(S_ID, pc_hair_hcc_t0))

write.xlsx(ss_data_hc, "2_OutputData/Hair_Cortisol_Data.xlsx")

# From prenatal data, we will use only CPSQ, RACD, and HC, as other metrics were duplicated at the 6-month visit. 
ss_data_0 <- ss_data %>%
  select(c(S_ID, ends_with("t0"))) %>%
  rename_with(~ str_remove(., "_t0")) %>%
  select(S_ID, starts_with("cpsq"), starts_with("racd"))

# Remove summary variables and ensure that data are numeric. 
ss_data_0_forimp <- ss_data_0 %>%
  select(-c(cpsq_mean:cpsq_anx, 
            racd_all:racdin)) %>%
  mutate(across(cpsq1:racd13, \(x) as.numeric(x)))

# Repeat for 6-month data
ss_data_6 <- ss_data %>%
  select(c(S_ID, ends_with("t6"))) %>%
  rename_with(~ str_remove(., "_t6")) %>%
  rename_with(.cols = c(adults_hh:income2needs_cat), .fn = function(.x) {paste0("hhi_", .x)}) %>%
  select(S_ID, starts_with("psi"), starts_with("bsi"), starts_with("esq"), starts_with("cts"), starts_with("crs"), starts_with("hhi"))

ss_data_6_forimp <- ss_data_6 %>%
  select(-c(psi_mn:psi_in, bsi_dep:bsi_psdi, esq_all:esq_in, cts_rv_agg:cts_in, 
            crs_agree:crs_in, hhi_adults_hh:hhi_total_hh, hhi_income2needs_ge1:hhi_income2needs_cat)) %>%
  mutate(across(psi1:crs16r, \(x) as.numeric(x)))
  
# Merge data frame back together
ss_data_forimp <- left_join(ss_data_0_forimp, ss_data_6_forimp, by = "S_ID") %>%
  column_to_rownames("S_ID")
```

Write out raw data for Dataverse:
```{r}
write.xlsx(ss_data_forimp %>%
  rownames_to_column("S_ID"), "2_OutputData/SS_Data_Raw.xlsx")
```


## Impute data with missForest

```{r warning = FALSE}
set.seed(946)
ss_data_imp <- missForest(ss_data_forimp)$ximp
```

## Recalculate summary variables
```{r}
ss_data_imp_summ_vars <- ss_data_imp %>%
  
  # Round data
  mutate_if(is.numeric, round, 4) %>%
  
  # CPSQ
  select(-c(cpsq4, cpsq5, cpsq7, cpsq8)) %>%
  mutate(cpsq_mean = rowMeans(across(cpsq1:cpsq10))) %>%
  mutate(cpsq_help = rowMeans(across(c(cpsq1, cpsq2, cpsq3, cpsq6, cpsq9, cpsq10)))) %>%
  mutate(cpsq_anx = rowMeans(across(c(cpsq4r, cpsq5r, cpsq7r, cpsq8r)))) %>%
  
  # RACD
  mutate(racd_all = rowMeans(across(racd1:racd13))) %>%
  mutate(racd_nozero = rowSums(across(racd1:racd13) != 0)) %>%
  
  # PSI
  select(-c(psi22, psi32, psi33)) %>%
  mutate(psi_mn = rowMeans(across(psi1:psi36))) %>%
  mutate(psi_pd = rowMeans(across(psi1:psi12))) %>%
  mutate(psi_pcdi = rowMeans(across(psi13:psi24))) %>%
  mutate(psi_dc = rowMeans(across(psi25:psi36))) %>%
  
  # BSI 
  mutate(bsi_som = rowMeans(across(c(bsi1, bsi4, bsi7, bsi10, bsi13, bsi16)))) %>%
  mutate(bsi_dep = rowMeans(across(c(bsi2, bsi5, bsi8, bsi11, bsi14, bsi17)))) %>%
  mutate(bsi_anx = rowMeans(across(c(bsi3, bsi6, bsi9, bsi12, bsi15, bsi18)))) %>%
  mutate(bsi_hos = rowMeans(across(c(bsi19, bsi20, bsi21, bsi22, bsi23)))) %>%
  mutate(bsi_gsi = rowMeans(across(bsi1:bsi23))) %>%
  mutate(bsi_psp = rowSums(across(bsi1:bsi23) != 0)/23) %>%
  mutate(across(bsi1:bsi23, ~na_if(., 0))) %>%
  mutate(bsi_psdi =  rowMeans(across(bsi1:bsi23), na.rm = TRUE)) %>%
  mutate(bsi_psdi = replace(bsi_psdi, is.nan(bsi_psdi), 0)) %>%
  mutate(bsi_psdi_psp = ifelse(bsi_psp != 0, bsi_psdi/bsi_psp, 0)) %>%
  mutate(across(bsi1:bsi23, ~replace_na(., 0))) %>%
  
  # ESQ
  mutate(esq_bill_scale = esq_bill*0.75) %>%
  mutate(esq_end_scale = esq_end*0.75) %>%
  mutate(esq_all = rowMeans(across(c(esq_bill_scale, esq_end_scale, esq_home:esq_med_ca)))) %>%
  select(-c(esq_bill_scale, esq_end_scale)) %>%
  
  # CTS
  mutate(cts_rv_agg = rowMeans(across(cts4a:cts10a))) %>%
  mutate(cts_pv_agg = rowMeans(across(cts4b:cts10b))) %>%
  mutate(cts_ev_agg = rowMeans(across(cts4c:cts10c))) %>%
  mutate(cts_rv_viol = rowMeans(across(cts11a:cts19a))) %>%
  mutate(cts_pv_viol = rowMeans(across(cts11b:cts19b))) %>%
  mutate(cts_ev_viol = rowMeans(across(cts11c:cts19c))) %>%
  
  # CRS
  select(-c(crs1, crs2, crs6, crs8, crs9, crs10, crs11, crs15, crs16)) %>%
  mutate(crs_mean = rowMeans(across(c(crs3, crs4, crs5, crs6r, crs7, crs8r, crs9r, crs10r, crs11r, crs12, crs13, crs14, crs15r, crs16r))))
```

## Detect outliers

```{r warning = FALSE}
# Filter and scale data frame
ss_data_for_clustering <- ss_data_imp_summ_vars %>% scale() %>% data.frame() %>% select_if(~ !any(is.na(.)))

# Run PCA
pca <- prcomp(ss_data_for_clustering)

# Create a scoring funciton to detect PCA sample outliers. The input is PCA results data frame and the number of standard deviations for the cutoff. The output is outlier names. 
outlier_detection = function(pca_df, sd){

    # getting scores
    scores = pca_df$x
    
    # identifying samples that are > 6 standard deviations away from the mean
    outlier_indices = apply(scores, 2, function(x) which( abs(x - mean(x)) > (sd * sd(x)) )) %>%
        Reduce(union, .)
    
    # getting sample names
    outliers = rownames(scores)[outlier_indices]
    
    return(outliers)
}

# Call function with different standard deviation cutoffs
outliers_6 <- outlier_detection(pca, 6)

ss_for_pca <- ss_data_for_clustering %>%
  rownames_to_column("S_ID") %>%
  mutate("Outlier Cutoff" = ifelse(S_ID %in% outliers_6, "6 SD Outlier", "Not Outlier")) %>%
  column_to_rownames("S_ID") %>%
  mutate(`Outlier Cutoff` = fct_relevel(`Outlier Cutoff`, c ("Not Outlier", "6 SD Outlier")))

# Graph PCA with outliers
PCA_outliers_plot <- fviz_pca_ind(pca, 
             label = "none",
             habillage = ss_for_pca$`Outlier Cutoff`,
             palette = c("gray12", "#CC3333"),
             mean.point = FALSE,
             pointsize = 3) +
  scale_shape_manual(values = c(20, 20, 20, 20)) +
theme(axis.title = element_text(face = "bold", size = rel(1.1)),
      panel.border = element_rect(fill = NA, color = "black", size = 0.3),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_blank(),
      plot.title = element_text(hjust = 0.5))

PCA_outliers_plot

# Graph PCA without outlier point
ss_for_pca_outrem <- ss_data_for_clustering %>%
  rownames_to_column("S_ID") %>%
  filter(S_ID != 1169) %>%
  column_to_rownames("S_ID")

pca <- prcomp(ss_for_pca_outrem)

PCA_outliers_plot_outrem <- fviz_pca_ind(pca, 
             label = "none",
             mean.point = FALSE,
             pointsize = 1.75) +
  scale_shape_manual(values = c(20, 20, 20, 20)) +
theme(axis.title = element_text(face = "bold", size = rel(1.1)),
      panel.border = element_rect(fill = NA, color = "black", size = 0.3),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_blank(),
      plot.title = element_text(hjust = 0.5))

PCA_outliers_plot_outrem
```

Create final figure panel for supplement:
```{r}
ss_outlier_det <- PCA_outliers_plot + PCA_outliers_plot_outrem +
  plot_layout(widths = c(4, 3.9)) +
  plot_annotation(tag_levels = 'A') & 
  theme(plot.tag = element_text(face = "bold", size = 18))

pdf(file = "3_OutputFigs/SS_Outlier_Removal.pdf",
    height = 4.5, 
    width = 10)

ss_outlier_det

invisible(dev.off())

ss_outlier_det
```

Export data:
```{r}
# Filter outlier subject
ss_data_cleaned <- ss_data_imp_summ_vars %>%
  rownames_to_column("S_ID") 

write.xlsx(ss_data_cleaned, "2_OutputData/SS_Data_Cleaned.xlsx")
```